---
title: "Congress_Demographics_Analysis"
author: "Emin Allen"
date: "2024-09-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-libraries}
# Load necessary libraries
library(readr)
library(dplyr)
library(ggplot2)
```

```{r load-data}
# Load data from GitHub
data <- read_csv("https://raw.githubusercontent.com/Emin-NYC/congress-age-data/main/data_aging_congress.csv")

# Create relevant subset of the data
data_subset <- data[, c("congress", "chamber", "age_years", "party_code", "generation")]

# Rename columns
colnames(data_subset) <- c("Congress", "Chamber", "Age", "Party", "Generation")

# Transform columns
data_subset$Party <- ifelse(data_subset$Party == 100, "Democrat", 
                            ifelse(data_subset$Party == 200, "Republican", "Other"))
data_subset$Age <- round(data_subset$Age)
```

## Testing the hypothesis: Is Congress Older than Ever Before?

```{r analysis}
# Calculate average age by Congress
congress_average_age <- data_subset %>%
  group_by(Congress) %>%
  summarise(Average_Age = mean(Age, na.rm = TRUE))

# Show the first few rows of the result
head(congress_average_age)
```

## Plotting a graph to visualize the average age of Congress over time


```{r plot-age}
# Plot average age over time
ggplot(congress_average_age, aes(x = Congress, y = Average_Age)) +
  geom_line(color = "blue") +
  geom_point() +
  labs(title = "Average Age of Congress over Time", x = "Congress", y = "Average Age") +
  theme_minimal()
```

## Conclusion

Our graph results show that the average age of Congress has indeed increased over time, confirming the hypothesis. Future analysis could potentially explore the differences in age between political parties and how that might influence policy making.